---
title: "Assignment 4 - Applying meta-analytic priors"
author: "Riccardo Fusaroli"
date: "3/16/2018"
output: html_document
---

## Assignment 4

In this assignment we do the following:
- we reproduce the meta-analysis of pitch SD from last semester in a Bayesian framework
- we reproduce the pitch SD in schizophrenia analysis from last semester using both a conservative and a meta-analytic prior
- we assess the difference in model quality and estimates using the two priors.

The questions you need to answer are: What are the consequences of using a meta-analytic prior? Evaluate the models with conservative and meta-analytic priors. Discuss the effects on estimates. Discuss the effects on model quality. Discuss the role that meta-analytic priors should have in scientific practice. Should we systematically use them? Do they have drawbacks? Should we use them to complement more conservative approaches? How does the use of meta-analytic priors you suggest reflect the skeptical and cumulative nature of science?

### Step by step suggestions

Step 1: Reproduce the meta-analysis of pitch sd from previous studies of voice in schizophrenia
- the data is available as Assignment4MetaData.xlsx

```{r load data}
meta = readxl::read_excel("Assignment4MetaData.xlsx")
```

- Effect size (cohen's d), sd and variance are already calculated (you're welcome!)
- Since we're only interested in getting a meta-analytic effect size, let's take a shortcut and use bromance magic (brms): https://mvuorre.github.io/post/2016/2016-09-29-bayesian-meta-analysis/
```{r libraries}
library(brms)
library(metafor)
library(lme4)
library(tidyverse)
library(rethinking)
library(bayesplot)

#devtools::install_github("mvuorre/brmstools")
library(brmstools)
library(ggplot2)
library(ggthemes)
```

```{r clean data}
#keep only the variables we need to fit the meta-effect model
data = meta[,c(5,29,31)]
data = na.omit(data)
```

```{r}
meta_effects <- brm(MeanES | se(SdES) ~ 1 + (1|StudyRef),
               data = data, iter = 20000, cores = 2)

summary(meta_effects)

#save the output of the above model for later use
meta_prior = alist(dnorm(-0.54, 0.24))

#forest plot
forest(meta_effects,
       show_data = TRUE,
       av_name = "Effect size")+
  theme_few(base_family = "Verdana")+
  labs(title="Estimated effects of all\nstudies and meta-effect",x="")

dev.copy(jpeg,'Figures/forest_plot.jpg')
dev.off()
```


Step 2: Prepare the pitch SD data from last year
- the data is available as Assignment4PitchData.csv (thanks Celine)
- We do not know how to build random effects, yet. So we need to simplify the dataset to avoid interdependence between datapoint: How?
- Also, let's standardize the data, so that they are compatible with our meta-analytic prior (Cohen's d is measured in SDs).

```{r load data}
library(magrittr)
pitch = readxl::read_excel("Assignment4PitchDatav2.xlsx")

pitch_sum = pitch %>% 
  group_by(ID_unique) %>%
  mutate(SD=mean(PitchSD)) %>%
  ungroup() %>%
  select(1:3,5,13)

pitch_ready=pitch_sum[!duplicated(pitch_sum),]
pitch_ready$SD_s = (pitch_ready$SD-mean(pitch_ready$SD))/sd(pitch_ready$SD)

mean(pitch_ready$SD_s)
sd(pitch_ready$SD_s)
```


Step 3: Build a regression model predicting Pitch SD from Diagnosis.
- how is the outcome distributed? (likelihood function)
- how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on?
- use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it.
- Describe and plot the estimates. Evaluate model quality

```{r}
library(ggplot2)

data_m2s = pitch_ready %>% select(2,6) %>% as.data.frame()


ggplot(data_m2s,aes(SD_s))+
  geom_line(stat = "density")+
  facet_wrap("diagnosis")+
  theme_few()



model_skep = map2stan( 
  flist = alist(
    SD_s ~ dnorm(mu,sigma),
    mu <- a + bD*diagnosis ,
    a ~ dnorm( 0, 1) ,
    bD ~ dnorm( 0 ,0.2 ) ,
    sigma ~ dcauchy(0 ,1)),
  data=data_m2s,
  iter=20000,
  chains = 2,
  cores = 2
)

plot(model_skep)

pairs(model_skep)
dev.copy(jpeg,'Figures/Pairs_skeptical.jpg')
dev.off()

(precis_skep=precis(model_skep))
write.csv(precis_skep@output,"Results/Skeptical_output.csv")
```

```{r plotting}
#PPC plot
library(bayesplot)

post_skep = sim(model_skep,n=1e4)
y=data_m2s$SD_s

PPC_skep=pp_check(y, post_skep[1:300,],"dens_overlay")+
  theme_few(base_family = "Verdana")+
  labs(title="PPC plot of model with skeptical prior",x="Standard Deviation" , y= "Density")+
    scale_color_manual(name="Distributions",
                         breaks=c("y","yrep"),
                         labels=c("Training", "Prediction"),
                       values = c("black","steelblue"))

ggsave("Figures/PPC_Skeptical.jpg", PPC_skep)
```

```{r plot results}
mu = link(model_skep,n=1e4)
mu_mean = apply( mu , 2 , mean )
mu_PI = as.data.frame(t(apply( mu , 2 , PI , prob=0.89 )))
SD_sim = sim(model_skep,n=1e4)
SD_PI = as.data.frame(t(apply( outcome_sim , 2 , PI , prob=0.89 )))

skep_results = ggplot(data_m2s)+
  #plot the real data
  geom_point(aes(x=diagnosis,y=SD_s))+
  #plot predicted mean
  geom_line(data=data.frame(x=data_m2s$diagnosis, y=mu_mean),aes(x=x,y=y))+
  #plot uncertainty around predicted mean
  geom_ribbon(data=mu_PI,aes(ymin=`5%`,ymax=`94%`,x=data_m2s$diagnosis),alpha=0.6,fill="blue")+
  geom_ribbon(data=SD_PI,aes(ymin=`5%`,ymax=`94%`,x=data_m2s$diagnosis),alpha=0.5,fill="blue")+
  labs(title="Regression of SD from Diagnosis",x="Diagnosis",y="Standard Deviation")+
  scale_x_continuous(breaks=c(0,1),
                     labels=c("Control","Schizo"))+
  theme_few()

skep_results

ggsave("Figures/Skeptical_results.jpg",skep_results)
```


```{r}
model_skep2 = map2stan( 
  flist = alist(
    SD_s ~ dnorm(mu,sigma),
    mu <- a + bD*diagnosis ,
    a ~ dnorm( 0, 1) ,
    bD ~ dnorm( 0 , 0.15 ) ,
    log(sigma) ~ a_sigma + bD_sigma*diagnosis,
    a_sigma ~dnorm(0,1),
    bD_sigma ~ dnorm(0,0.1)
      ),
  data=data_m2s,
  iter=10000,
  chains=2,
  cores=2
)

precis(model_skep2)
plot(model_skep2)
pairs(model_skep2)
```

```{r}
compare(model_skep,model_skep2) #the simple one is better
```



Step 4: Now re-run the model with the meta-analytic prior
- Describe and plot the estimates. Evaluate model quality

```{r}
model_meta = map2stan( 
  flist = alist(
    SD_s ~ dnorm(mu,sigma),
    mu <- a + bD*diagnosis ,
    a ~ dnorm( 0, 1) ,
    bD ~ dnorm(-0.55, 0.25) ,
    sigma ~ dcauchy(0 ,1)),
  data=data_m2s,
  iter=20000,
  chains = 2,
  cores = 2
)

plot(model_meta)

pairs(model_meta)
dev.copy(jpeg,'Figures/Pairs_meta.jpg')
dev.off()

(precis_meta=precis(model_meta))
write.csv(precis_meta@output,"Results/Meta_output.csv")
```

```{r}
post_meta = sim(model_meta,n=1e4)
y=data_m2s$SD_s

(meta_PPC=pp_check(y, post_meta[1:200,],"dens_overlay")+
  theme_few(base_family = "Verdana")+
  labs(title="PPC plot of model with meta-analytic prior",x="Standard Deviation" , y= "Density")+
    scale_color_manual(name="Distributions",
                         breaks=c("y","yrep"),
                         labels=c("Training", "Prediction"),
                       values = c("black","steelblue"))
    
  )

ggsave("Figures/PPC_Meta.jpg",meta_PPC)
```


Step 5: Compare the models
- Plot priors and posteriors of the diagnosis effect in both models
```{r}
#draw data from priors
skep_prior = rnorm(1e4,0,0.15)
meta_prior = rnorm(1e4,-0.55, 0.25)

#draw from posteriors of the diagnosis effect
post_skep = extract.samples(model_skep,n=1e4)
post_meta = extract.samples(model_meta,n=1e4)

dia_skep = post_skep$bD
dia_meta = post_meta$bD

#put the draws together for plotting
prior_post = data.frame(skep_prior,dia_skep,meta_prior,dia_meta)

melt_prior_post = reshape2::melt(prior_post)

library(ggthemes)
prior_posterior=ggplot(melt_prior_post,aes(value,colour=variable))+
  geom_line(stat = "density")+
  labs(title="Priors and posteriors of the Diagnosis effect", subtitle="SM = Skeptical Model\nMM = Meta-analytic Model",x="SD" , y= "Density")+
  theme_few(base_family = "Verdana")+
  scale_color_manual(name="Distributions",
                       breaks=c("skep_prior","dia_skep","meta_prior","dia_meta"),
                       labels=c("SM Prior","SM Posterior", "MM Prior", "MM Posterior"),
                     values = c("#e50000","#660000", '#b2d8b2', "#008000"))

prior_posterior
ggsave("Figures/Priors_&_Posteriors.jpg",prior_posterior)
```

- Compare posteriors between the two models
```{r}
posteriors_melt = subset(melt_prior_post,variable==c("dia_skep","dia_meta"))

(posteriors_pl=ggplot(posteriors_melt,aes(value,colour=variable))+
  geom_line(stat = "density")+
  labs(title="Posteriors of the Diagnosis effect",x="SD" , y= "Density")+
  theme_few(base_family = "Verdana")+
  scale_color_discrete(name="Posteriors",
                       breaks=c('dia_skep',"dia_meta"),
                       labels=c('Skeptical Posterior', "Meta Posterior"))
  )

ggsave("Figures/Posteriors_compare.jpg", posteriors_pl)

#compute the probability that there is meta-posterior is larger
diff=dia_meta-dia_skep
(prob=sum(diff<0)/1e4)

#or that the difference is reversed
1-prob
```

- Compare their relative distance from truth (WAIC)
```{r}
(WAIC_compare=compare(model_skep,model_meta))

#save the output
write.csv(WAIC_compare@output,"Results/WAIC_compare.csv")

co_tab=coeftab(model_skep,model_meta)
coeftab_plot(co_tab)

dev.copy(jpeg,'Figures/Compare_coefficients.jpg')
dev.off()
```

EXTRA:
Construct an ensemble model to predict the data. Make PPC plot.
```{r}
ensemble=ensemble(model_skep,model_meta,n=1e4)
ens_pred = ensemble$sim
y=data_m2s$SD_s
pp_check(y, ens_pred[1:300,],"dens_overlay")
```

- Discuss how they compare and whether any of them is best.

Step 6: Prepare a nice write up of the analysis and answer the questions at the top.

Optional step 7: how skeptical should a prior be?
- Try different levels of skepticism and compare them using WAIC.

Optional step 8: Include other predictors
- Do age, gender and education improve the model?
- Should they be main effects or interactions?

Optional step 9: Bromance magic.
- explore the bromance code below including random effects (by default with weakly informative priors)
- learn how to change the prior
- explore effects of trial, age, gender, including the appropriate random slopes
- compare the models you created using WAIC and posterior predictive check (pp_check())


```{r}

brm_out <- brm(PitchSD ~ 1 + Diagnosis  +(1|ID_unique/Study), # Outcome as a function of the predictors as in lme4. 
               data=pitch, # Define the data
               family=gaussian(), # Define the family. 
               iter = 5000, warmup = 2000, cores = 4)
summary(brm_out1)
plot(brm_out1)

```

